---
output:
  pdf_document: default
  html_document: default
---
# R-codes

## Import dataset

```{r}
heart <- read.csv("heart.csv")
attach(heart)
head(heart)
```

## Analyse variables

```{r}
#install.packages("ggpubr")
library(ggpubr)
library(ggplot2)
```


### 1) age - age - Numerical

```{r}
table(age)
age_hist = ggplot(data = heart, aes(x = age, fill = factor(target))) +
  geom_histogram(color="black", alpha=0.5, position="identity",bins=15) +
  labs(title = "age",
       x = "age",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()


```

### 2) sex - sex - Categorical

```{r}
table(sex)
sex_bar = ggplot(data = heart, aes(x = factor(sex), fill = factor(target))) +
  geom_bar(color="black", alpha=0.5) +
  labs(title = "sex",
       x = "sex",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()

```

### 3) cp - chest pain type (4 values) - Categorical

```{r}
table(cp)
cp_bar = ggplot(data = heart, aes(x = factor(cp), fill = factor(target))) +
  geom_bar(color="black", alpha=0.5) +
  labs(title = "cp",
       x = "cp",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()

```

### 4) trestbps - resting blood pressure - Numerical

```{r}
table(trestbps)

trestbps_hist = ggplot(data = heart, aes(x = trestbps, fill = factor(target))) +
  geom_histogram(color="black", alpha=0.5, position="identity",bins=15) +
  labs(title = "trestbps",
       x = "trestbps",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()


```

### 5) chol - serum cholestoral in mg/dl - Numerical

```{r}
table(chol)

chol_hist = ggplot(data = heart, aes(x = chol, fill = factor(target))) +
  geom_histogram(color="black", alpha=0.5, position="identity",bins=15) +
  labs(title = "chol",
       x = "chol",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()

```

### 6) fbs - fasting blood sugar \> 120 mg/dl - Categorical

```{r}
table(fbs)
fbs_bar = ggplot(data = heart, aes(x = factor(fbs), fill = factor(target))) +
  geom_bar(color="black", alpha=0.5) +
  labs(title = "fbs",
       x = "fbs",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()
```

### 7) restecg - resting electrocardiographic results (values 0,1,2) - Categorical

```{r}
table(restecg)
restecg_bar = ggplot(data = heart, aes(x = factor(restecg), fill = factor(target))) +
  geom_bar(color="black", alpha=0.5) +
  labs(title = "restecg",
       x = "restecg",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()
```

### 8) thalach - maximum heart rate achieved - Numerical

```{r}
table(thalach)

thalach_hist = ggplot(data = heart, aes(x = thalach, fill = factor(target))) +
  geom_histogram(color="black", alpha=0.5, position="identity",bins=15) +
  labs(title = "thalach",
       x = "thalach",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()
```

### 9) exang - exercise induced angina - Categorical

```{r}
table(exang)
exang_bar = ggplot(data = heart, aes(x = factor(exang), fill = factor(target))) +
  geom_bar(color="black", alpha=0.5) +
  labs(title = "exang",
       x = "exang",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()
```

### 10) oldpeak - ST depression induced by exercise relative to rest - Numerical

```{r}
table(oldpeak)

oldpeak_hist = ggplot(data = heart, aes(x = oldpeak, fill = factor(target))) +
  geom_histogram(color="black", alpha=0.5, position="identity",bins=15) +
  labs(title = "oldpeak",
       x = "oldpeak",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()

```

### 11) slope- the slope of the peak exercise ST segment - Categorical

```{r}
table(slope)
slope_bar = ggplot(data = heart, aes(x = factor(slope), fill = factor(target))) +
  geom_bar(color="black", alpha=0.5) +
  labs(title = "slope",
       x = "slope",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()
```

### 12) ca - number of major vessels (0-3) colored by flourosopy - Categorical

```{r}
table(ca)
ca_bar = ggplot(data = heart, aes(x = factor(ca), fill = factor(target))) +
  geom_bar(color="black", alpha=0.5) +
  labs(title = "ca",
       x = "ca",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()
```

### 13) Target (y)

```{r}
table(target)

ggplot(data = heart, aes(x = factor(target), fill = factor(target))) +
  geom_bar(color="black", alpha=0.5) +
  labs(title = "target",
       x = "target",
       y = "Count") +
  scale_fill_manual(values = c("aquamarine2", "mediumpurple2"), labels = c("0", "1"), name = "Target") +
  theme_minimal()


```

### Categorical plots

```{r}
categorical_plots <- ggarrange(sex_bar, cp_bar, fbs_bar, restecg_bar,exang_bar, slope_bar, ca_bar, 
                    #labels = c("sex", "cp", "fbs","restecg","exang","slope","ca"),
                    ncol = 3, nrow = 3)
categorical_plots
```

### Numerical plots

```{r}
numerical_plots <- ggarrange(age_hist, trestbps_hist, chol_hist, thalach_hist, oldpeak_hist,
                    ncol = 3, nrow = 2)

numerical_plots
```

## Multicollinearity for numerical variables

```{r}
heart_num = data.frame(age,trestbps,chol,thalach,oldpeak)
cor(heart_num)
```

## Purposeful Selection

### 1. Construct an initial main-effects model using explanatory variables that include the known important variables and others that show any evidence of being relevant when used as sole predictors (e.g., having P -value \< 0.2).

```{r}
#Create the base model 
model_base <- glm(target ~ 1, family=binomial, data=heart)
summary(model_base)
```

```{r}

#1 - age - age - Numerical
model_age <- glm(target ~ age, family = binomial, data=heart)
a_age = anova(model_base,model_age,test='LRT')

#2 - sex - sex - Categorical
model_sex <- glm(target ~ sex, family = binomial, data=heart)
a_sex = anova(model_base,model_sex,test='LRT')

#2 - cp - chest pain type (4 values) - Categorical
model_cp <- glm(target ~ factor(cp), family = binomial, data=heart)
a_cp = anova(model_base,model_cp,test='LRT')

#4 - trestbps - resting blood pressure - Numerical
model_trestbps <- glm(target ~ trestbps, family = binomial, data=heart)
a_trestbps = anova(model_base,model_trestbps,test='LRT')

#5 - chol - serum cholestoral in mg/dl - Numerical
model_chol <- glm(target ~ chol, family = binomial, data=heart)
a_chol = anova(model_base,model_chol,test='LRT')

#6 - fbs - fasting blood sugar \> 120 mg/dl - Categorical
model_fbs <- glm(target ~ factor(fbs), family = binomial, data=heart)
a_fbs = anova(model_base,model_fbs,test='LRT')

#7 - restecg - resting electrocardiographic results (values 0,1,2) - Categorical
model_restecg <- glm(target ~ factor(restecg), family = binomial, data=heart)
a_restecg = anova(model_base,model_restecg,test='LRT')

#8 - thalach - maximum heart rate achieved  - Numerical 
model_thalach <- glm(target ~ thalach, family = binomial, data=heart)
a_thalach = anova(model_base,model_thalach,test='LRT')

#9 - exang - exercise induced angina - Categorical
model_exang <- glm(target ~ factor(exang), family = binomial, data=heart)
a_exang = anova(model_base,model_exang,test='LRT')

#10 - oldpeak - ST depression induced by exercise relative to rest - Numerical 
model_oldpeak <- glm(target ~ oldpeak, family = binomial, data=heart)
a_oldpeak = anova(model_base,model_oldpeak,test='LRT')

#11 - slope- the slope of the peak exercise ST segment - Categorical 
model_slope <- glm(target ~ factor(slope), family = binomial, data=heart)
a_slope = anova(model_base,model_slope,test='LRT')

#12 - ca - number of major vessels (0-3) colored by flourosopy - Categorical
model_ca <- glm(target ~ factor(ca), family = binomial, data=heart)
a_ca = anova(model_base,model_ca,test='LRT')


```

#### Create table step 1

```{r}
alpha <- 0.2

Model = c(0,1,2,3,4,5,6,7,8,9,10,11,12)
ExplanatoryVariables = c('None','ag','se','cp','tr','ch','fb','re','th','ex','ol','sl','ca')
Deviance = c(model_base$deviance,
              a_age$`Resid. Dev`[2],
              a_sex$`Resid. Dev`[2],
              a_cp$`Resid. Dev`[2],
              a_trestbps$`Resid. Dev`[2],
              a_chol$`Resid. Dev`[2],
              a_fbs$`Resid. Dev`[2],
              a_restecg$`Resid. Dev`[2],
              a_thalach$`Resid. Dev`[2],
              a_exang$`Resid. Dev`[2],
              a_oldpeak$`Resid. Dev`[2],
              a_slope$`Resid. Dev`[2],
              a_ca$`Resid. Dev`[2]
             )
df = c(model_base$df.residual,
    a_age$`Resid. Df`[2],
    a_sex$`Resid. Df`[2],
    a_cp$`Resid. Df`[2],
    a_trestbps$`Resid. Df`[2],
    a_chol$`Resid. Df`[2],
    a_fbs$`Resid. Df`[2],
    a_restecg$`Resid. Df`[2],
    a_thalach$`Resid. Df`[2],
    a_exang$`Resid. Df`[2],
    a_oldpeak$`Resid. Df`[2],
    a_slope$`Resid. Df`[2],
    a_ca$`Resid. Df`[2]
)
dfDifference = c(NA,
    a_age$Df[2],
    a_sex$Df[2],
    a_cp$Df[2],
    a_trestbps$Df[2],
    a_chol$Df[2],
    a_fbs$Df[2],
    a_restecg$Df[2],
    a_thalach$Df[2],
    a_exang$Df[2],
    a_oldpeak$Df[2],
    a_slope$Df[2],
    a_ca$Df[2]
   )
AIC = c(NA,
    model_age$aic,
    model_sex$aic,
    model_cp$aic,
    model_trestbps$aic,
    model_chol$aic,
    model_fbs$aic,
    model_restecg$aic,
    model_thalach$aic,
    model_exang$aic,
    model_oldpeak$aic,
    model_slope$aic,
    model_ca$aic
   )
DevianceDifference = c(NA,
              a_age$Deviance[2],
              a_sex$Deviance[2],
              a_cp$Deviance[2],
              a_trestbps$Deviance[2],
              a_chol$Deviance[2],
              a_fbs$Deviance[2],
              a_restecg$Deviance[2],
              a_thalach$Deviance[2],
              a_exang$Deviance[2],
              a_oldpeak$Deviance[2],
              a_slope$Deviance[2],
              a_ca$Deviance[2]
             )
Pvalue = c(NA,
    a_age$`Pr(>Chi)`[2],
    a_sex$`Pr(>Chi)`[2],
    a_cp$`Pr(>Chi)`[2],
    a_trestbps$`Pr(>Chi)`[2],
    a_chol$`Pr(>Chi)`[2],
    a_fbs$`Pr(>Chi)`[2],
    a_restecg$`Pr(>Chi)`[2],
    a_thalach$`Pr(>Chi)`[2],
    a_exang$`Pr(>Chi)`[2],
    a_oldpeak$`Pr(>Chi)`[2],
    a_slope$`Pr(>Chi)`[2],
    a_ca$`Pr(>Chi)`[2]
)

SIGNIFICANT = c(NA,
    ifelse(a_age$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_sex$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_cp$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_trestbps$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_chol$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_fbs$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_restecg$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_thalach$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_exang$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_oldpeak$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_slope$`Pr(>Chi)`[2]< alpha,'YES','NO'),
    ifelse(a_ca$`Pr(>Chi)`[2]< alpha,'YES','NO')
)

ModelsCompared = c(NA,
    '(1)-(0)',
    '(2)-(0)',
    '(3)-(0)',
    '(4)-(0)',
    '(5)-(0)',
    '(6)-(0)',
    '(7)-(0)',
    '(8)-(0)',
    '(9)-(0)',
    '(10)-(0)',
    '(11)-(0)',
    '(12)-(0)')
step1 = data.frame(
  Model,ExplanatoryVariables,Deviance,df,AIC,ModelsCompared,DevianceDifference,dfDifference,Pvalue,SIGNIFICANT
)

step1


```

### 2. Conduct backward elimination, keeping a variable if it is either significant at a somewhat more stringent level or shows evidence of being a relevant confounder, in the sense that the estimated effect of a key variable changes substantially when it is removed.

Variables codes:

-   ag - age - Numerical

-   se - sex - Categorical

-   cp - cp - Categorical

-   tr - trestbps - Numerical

-   ch - chol - Numerical

-   fb - fbs - Categorical

-   re - restecg - Categorical

-   th - thalach - Numerical

-   ex - exang - Categorical

-   ol - oldpeak - Numerical

-   sl - slope - Categorical

-   ca - ca - Categorical

```{r}
#new model with the significant variables from step1
model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca <- glm(target ~ age+factor(sex)+factor(cp)+trestbps+chol+factor(restecg)+thalach+factor(exang)+oldpeak+factor(slope)+factor(ca), family = binomial, data=heart)
step1_base = anova(model_base,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP ag - age - Numerical
model_se.cp.tr.ch.re.th.ex.ol.sl.ca <- glm(target ~ factor(sex)+factor(cp)+trestbps+chol+factor(restecg)+thalach+factor(exang)+oldpeak+factor(slope)+factor(ca),
                                           family = binomial, data=heart)
drop_age = anova(model_se.cp.tr.ch.re.th.ex.ol.sl.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP se - sex - Categorical
model_ag.cp.tr.ch.re.th.ex.ol.sl.ca <- glm(target ~ age+factor(cp)+trestbps+chol+factor(restecg)+thalach+factor(exang)+oldpeak+factor(slope)+factor(ca), 
                                           family = binomial, data=heart)
drop_sex = anova(model_ag.cp.tr.ch.re.th.ex.ol.sl.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP cp - cp - Categorical
model_ag.se.tr.ch.re.th.ex.ol.sl.ca <- glm(target ~ age+factor(sex)+trestbps+chol+factor(restecg)+thalach+factor(exang)+oldpeak+factor(slope)+factor(ca), 
                                           family = binomial, data=heart)
drop_cp = anova(model_ag.se.tr.ch.re.th.ex.ol.sl.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP tr - trestbps - Numerical
model_ag.se.cp.ch.re.th.ex.ol.sl.ca <- glm(target ~ age+factor(sex)+factor(cp)+chol+factor(restecg)+thalach+factor(exang)+oldpeak+factor(slope)+factor(ca), 
                                           family = binomial, data=heart)
drop_trestbps = anova(model_ag.se.cp.ch.re.th.ex.ol.sl.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP ch - chol - Numerical
model_ag.se.cp.tr.re.th.ex.ol.sl.ca <- glm(target ~ age+factor(sex)+factor(cp)+trestbps+factor(restecg)+thalach+factor(exang)+oldpeak+factor(slope)+factor(ca), 
                                           family = binomial, data=heart)
drop_chol = anova(model_ag.se.cp.tr.re.th.ex.ol.sl.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP re - restecg - Categorical
model_ag.se.cp.tr.ch.th.ex.ol.sl.ca <- glm(target ~ age+factor(sex)+factor(cp)+trestbps+chol+thalach+factor(exang)+oldpeak+factor(slope)+factor(ca), 
                                           family = binomial, data=heart)
drop_restecg = anova(model_ag.se.cp.tr.ch.th.ex.ol.sl.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP th - thalach - Numerical
model_ag.se.cp.tr.ch.re.ex.ol.sl.ca <- glm(target ~ age+factor(sex)+factor(cp)+trestbps+chol+factor(restecg)+factor(exang)+oldpeak+factor(slope)+factor(ca), 
                                           family = binomial, data=heart)
drop_thalach = anova(model_ag.se.cp.tr.ch.re.ex.ol.sl.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP ex - exang - Categorical
model_ag.se.cp.tr.ch.re.th.ol.sl.ca <- glm(target ~ age+factor(sex)+factor(cp)+trestbps+chol+factor(restecg)+thalach+oldpeak+factor(slope)+factor(ca), 
                                           family = binomial, data=heart)
drop_exang = anova(model_ag.se.cp.tr.ch.re.th.ol.sl.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP ol - oldpeak - Numerical
model_ag.se.cp.tr.ch.re.th.ex.sl.ca <- glm(target ~ age+factor(sex)+factor(cp)+trestbps+chol+factor(restecg)+thalach+factor(exang)+factor(slope)+factor(ca), 
                                           family = binomial, data=heart)
drop_oldpeak = anova(model_ag.se.cp.tr.ch.re.th.ex.sl.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP sl - slope - Numerical
model_ag.se.cp.tr.ch.re.th.ex.ol.ca <- glm(target ~ age+factor(sex)+factor(cp)+trestbps+chol+factor(restecg)+thalach+factor(exang)+oldpeak+factor(ca), 
                                           family = binomial, data=heart)
drop_slope = anova(model_ag.se.cp.tr.ch.re.th.ex.ol.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

#DROP ca - ca - Categorical
model_ag.se.cp.tr.ch.re.th.ex.ol.sl <- glm(target ~ age+factor(sex)+factor(cp)+trestbps+chol+factor(restecg)+thalach+factor(exang)+oldpeak+factor(slope), 
                                           family = binomial, data=heart)
drop_ca = anova(model_ag.se.cp.tr.ch.re.th.ex.ol.sl,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT')

```

#### Create table step 2

```{r}
alpha2 <- 0.05

Model2 = c(13,14,15,16,17,18,19,20,21,22,23,24)
ExplanatoryVariables2 = c(
    'ag+se+cp+tr+ch+re+th+ex+ol+sl+ca',
    'se+cp+tr+ch+re+th+ex+ol+sl+ca',
    'ag+cp+tr+ch+re+th+ex+ol+sl+ca',
    'ag+se+tr+ch+re+th+ex+ol+sl+ca',
    'ag+se+cp+ch+re+th+ex+ol+sl+ca',
    'ag+se+cp+tr+re+th+ex+ol+sl+ca',
    'ag+se+cp+tr+ch+th+ex+ol+sl+ca',
    'ag+se+cp+tr+ch+re+ex+ol+sl+ca',
    'ag+se+cp+tr+ch+re+th+ol+sl+ca',
    'ag+se+cp+tr+ch+re+th+ex+sl+ca',
    'ag+se+cp+tr+ch+re+th+ex+ol+ca',
    'ag+se+cp+tr+ch+re+th+ex+ol+sl'
)

Deviance2 = c(step1_base$`Resid. Dev`[1],
              drop_age$`Resid. Dev`[1],
              drop_sex$`Resid. Dev`[1],
              drop_cp$`Resid. Dev`[1],
              drop_trestbps$`Resid. Dev`[1],
              drop_chol$`Resid. Dev`[1],
              drop_restecg$`Resid. Dev`[1],
              drop_thalach$`Resid. Dev`[1],
              drop_exang$`Resid. Dev`[1],
              drop_oldpeak$`Resid. Dev`[1],
              drop_slope$`Resid. Dev`[1],
              drop_ca$`Resid. Dev`[1]
             )
df2 = c(step1_base$`Resid. Df`[1],
    drop_age$`Resid. Df`[1],
    drop_sex$`Resid. Df`[1],
    drop_cp$`Resid. Df`[1],
    drop_trestbps$`Resid. Df`[1],
    drop_chol$`Resid. Df`[1],
    drop_restecg$`Resid. Df`[1],
    drop_thalach$`Resid. Df`[1],
    drop_exang$`Resid. Df`[1],
    drop_oldpeak$`Resid. Df`[1],
    drop_slope$`Resid. Df`[1],
    drop_ca$`Resid. Df`[1]
)
dfDifference2 = c(step1_base$Df[2],
    drop_age$Df[2],
    drop_sex$Df[2],
    drop_cp$Df[2],
    drop_trestbps$Df[2],
    drop_chol$Df[2],
    drop_restecg$Df[2],
    drop_thalach$Df[2],
    drop_exang$Df[2],
    drop_oldpeak$Df[2],
    drop_slope$Df[2],
    drop_ca$Df[2]
   )
AIC2 = c(model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca$aic,
    model_se.cp.tr.ch.re.th.ex.ol.sl.ca$aic,
    model_ag.cp.tr.ch.re.th.ex.ol.sl.ca$aic,
    model_ag.se.tr.ch.re.th.ex.ol.sl.ca$aic,
    model_ag.se.cp.ch.re.th.ex.ol.sl.ca$aic,
    model_ag.se.cp.tr.re.th.ex.ol.sl.ca$aic,
    model_ag.se.cp.tr.ch.th.ex.ol.sl.ca$aic,
    model_ag.se.cp.tr.ch.re.ex.ol.sl.ca$aic,
    model_ag.se.cp.tr.ch.re.th.ol.sl.ca$aic,
    model_ag.se.cp.tr.ch.re.th.ex.sl.ca$aic,
    model_ag.se.cp.tr.ch.re.th.ex.ol.ca$aic,
    model_ag.se.cp.tr.ch.re.th.ex.ol.sl$aic
 )
DevianceDifference2 = c(step1_base$Deviance[2],
              drop_age$Deviance[2],
              drop_sex$Deviance[2],
              drop_cp$Deviance[2],
              drop_trestbps$Deviance[2],
              drop_chol$Deviance[2],
              drop_restecg$Deviance[2],
              drop_thalach$Deviance[2],
              drop_exang$Deviance[2],
              drop_oldpeak$Deviance[2],
              drop_slope$Deviance[2],
              drop_ca$Deviance[2]
             )
Pvalue2 = c(step1_base$`Pr(>Chi)`[2],
    drop_age$`Pr(>Chi)`[2],
    drop_sex$`Pr(>Chi)`[2],
    drop_cp$`Pr(>Chi)`[2],
    drop_trestbps$`Pr(>Chi)`[2],
    drop_chol$`Pr(>Chi)`[2],
    drop_restecg$`Pr(>Chi)`[2],
    drop_thalach$`Pr(>Chi)`[2],
    drop_exang$`Pr(>Chi)`[2],
    drop_oldpeak$`Pr(>Chi)`[2],
    drop_slope$`Pr(>Chi)`[2],
    drop_ca$`Pr(>Chi)`[2]
)

SIGNIFICANT2 = c(
    ifelse(step1_base$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_age$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_sex$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_cp$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_trestbps$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_chol$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_restecg$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_thalach$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_exang$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_oldpeak$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_slope$`Pr(>Chi)`[2]< alpha2,'YES','NO'),
    ifelse(drop_ca$`Pr(>Chi)`[2]< alpha2,'YES','NO')
)

ModelsCompared2 = c('(13)-(0)',
                  '(13)-(14)',
                  '(13)-(15)',
                  '(13)-(16)',
                  '(13)-(17)',
                  '(13)-(18)',
                  '(13)-(19)',
                  '(13)-(20)',
                  '(13)-(21)',
                  '(13)-(22)',
                  '(13)-(23)',
                  '(13)-(24)')
step2 = data.frame(
  Model2,ExplanatoryVariables2,Deviance2,df2,AIC2,ModelsCompared2,DevianceDifference2,dfDifference2,Pvalue2,SIGNIFICANT2
)

step2

```

### 3. Add to the model any variables that were not included in step1 but that are significant when adjusting for the variables in the model after step 2, since a variable may not be significantly associated with y but may make an important contribution in the presence of other variables.

-   se - sex - Categorical

-   cp - cp - Categorical

-   tr - trestbps - Numerical

-   ex - exang - Categorical

-   sl - slope - Numerical

-   ca - ca - Categorical

-   Add fb - fbs - Categorical

```{r}
#useful model until now
model_se.cp.tr.ex.sl.ca <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca), family = binomial, data=heart)
step3_model = anova(model_se.cp.tr.ex.sl.ca,model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca,test='LRT') #13

#add fbs
model_se.cp.tr.ex.sl.ca.fb <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(fbs), family = binomial, data=heart)
add_fbs = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca.fb,test='LRT')

```

#### Create table step 3

```{r}
alpha3 <- 0.05

Model3 = c(25,26)
ExplanatoryVariables3 = c('se+cp+tr+ex+sl+ca','se+cp+tr+ex+sl+ca+fb')
Deviance3 = c(NA,
              add_fbs$`Resid. Dev`[2]
             )
df3 = c(NA,
    add_fbs$`Resid. Df`[2]
)
dfDifference3 = c(NA,
    add_fbs$Df[2]
   )
AIC3 = c(model_ag.se.cp.tr.ch.re.th.ex.ol.sl.ca$aic,
    model_se.cp.tr.ex.sl.ca.fb$aic
   )
DevianceDifference3 = c(NA,
              add_fbs$Deviance[2]
             )

Pvalue3 = c(NA,
    add_fbs$`Pr(>Chi)`[2]
)

SIGNIFICANT3 = c(
    NA,
    ifelse(add_fbs$`Pr(>Chi)`[2]< alpha3,'YES','NO')
)

ModelsCompared3 = c('(25)-(13)',
                  '(26)-(25)')
step3 = data.frame(
  Model3,ExplanatoryVariables3,Deviance3,df3,AIC3,ModelsCompared3,DevianceDifference3,dfDifference3,Pvalue3,SIGNIFICANT3
)



```

```{r}
step3
```

### 4 Check for plausible interactions among variables in the model after step 3, using significance tests at conventional levels such as 0.05.

-   se - sex - Categorical

-   cp - cp - Categorical

-   tr - trestbps - Numerical

-   ex - exang - Categorical

-   sl - slope - Numerical

-   ca - ca - Categorical

```{r}
#useful model until now
model_se.cp.tr.ex.sl.ca <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca), family = binomial, data=heart)

# interaction between sex and cp
model_se.cp.tr.ex.sl.ca_se.cp <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(sex):factor(cp), family = binomial, data=heart)
int_sex_cp = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_se.cp,test='LRT')

# interaction between sex and trestbps
model_se.cp.tr.ex.sl.ca_se.tr <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(sex):trestbps, family = binomial, data=heart)
int_sex_trestbps = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_se.tr,test='LRT')

# interaction between sex and exang
model_se.cp.tr.ex.sl.ca_se.ex <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(sex):factor(exang), family = binomial, data=heart)
int_sex_exang = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_se.ex,test='LRT')

# interaction between sex and slope
model_se.cp.tr.ex.sl.ca_se.sl <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(sex):slope, family = binomial, data=heart)
int_sex_slope = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_se.sl,test='LRT')

# interaction between sex and ca
model_se.cp.tr.ex.sl.ca_se.ca <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(sex):factor(ca), family = binomial, data=heart)
int_sex_ca = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_se.ca,test='LRT')

```

```{r}

###cp

# interaction between cp and trestbps
model_se.cp.tr.ex.sl.ca_cp.tr <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(cp):trestbps, 
                                        family = binomial, data=heart)
int_cp_trestbps = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_cp.tr,test='LRT')

# interaction between cp and exang
model_se.cp.tr.ex.sl.ca_cp.ex <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(cp):factor(exang), 
                                        family = binomial, data=heart)
int_cp_exang = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_cp.ex,test='LRT')

# interaction between cp and slope
model_se.cp.tr.ex.sl.ca_cp.sl <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(cp):slope, 
                                        family = binomial, data=heart)
int_cp_slope = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_cp.sl,test='LRT')

# interaction between cp and ca
model_se.cp.tr.ex.sl.ca_cp.ca <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(cp):factor(ca), 
                                        family = binomial, data=heart)
int_cp_ca = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_cp.ca,test='LRT')
```

```{r}

###trestbps

# interaction between trestbps and exang
model_se.cp.tr.ex.sl.ca_tr.ex <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+trestbps:factor(exang), 
                                        family = binomial, data=heart)
int_trestbps_exang = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_tr.ex,test='LRT')

# interaction between trestbps and slope
model_se.cp.tr.ex.sl.ca_tr.sl <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+trestbps:slope, 
                                        family = binomial, data=heart)
int_trestbps_slope = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_tr.sl,test='LRT')

# interaction between trestbps and ca
model_se.cp.tr.ex.sl.ca_tr.ca <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+trestbps:factor(ca), 
                                        family = binomial, data=heart)
int_trestbps_ca = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_tr.ca,test='LRT')

```

```{r}

#exang

# interaction between exang and slope
model_se.cp.tr.ex.sl.ca_ex.sl <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(exang):slope, 
                                        family = binomial, data=heart)
int_exang_slope = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_ex.sl,test='LRT')

# interaction between exang and ca
model_se.cp.tr.ex.sl.ca_ex.ca <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(exang):factor(ca), 
                                     family = binomial, data=heart)
int_exang_ca = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_ex.ca,test='LRT')

```

```{r}

#slope

# interaction between slope and ca
model_se.cp.tr.ex.sl.ca_sl.ca <- glm(target ~ factor(sex)+factor(cp)+trestbps+factor(exang)+factor(slope)+factor(ca)+factor(slope):factor(ca), 
                                        family = binomial, data=heart)
int_slope_ca = anova(model_se.cp.tr.ex.sl.ca,model_se.cp.tr.ex.sl.ca_sl.ca,test='LRT')

```

#### Create table step 4

```{r}
alpha4 <- 0.05

Model4 = c(27,28,29,30,31,32,33,34,35,36,37,38,39,40,41)

ExplanatoryVariables4 = c(
    'se+cp+tr+ex+sl+ca+se*cp'
    ,'se+cp+tr+ex+sl+ca+se*tr'
    ,'se+cp+tr+ex+sl+ca+se*ex'
    ,'se+cp+tr+ex+sl+ca+se*sl'
    ,'se+cp+tr+ex+sl+ca+se*ca'
    ,'se+cp+tr+ex+sl+ca+cp*tr'
    ,'se+cp+tr+ex+sl+ca+cp*ex'
    ,'se+cp+tr+ex+sl+ca+cp*sl'
    ,'se+cp+tr+ex+sl+ca+cp*ca'
    ,'se+cp+tr+ex+sl+ca+tr*ex'
    ,'se+cp+tr+ex+sl+ca+tr*sl'
    ,'se+cp+tr+ex+sl+ca+tr*ca'
    ,'se+cp+tr+ex+sl+ca+ex*sl'
    ,'se+cp+tr+ex+sl+ca+ex*ca'
    ,'se+cp+tr+ex+sl+ca+sl*ca')


Deviance4 = c(
    int_sex_cp$`Resid. Dev`[2],
    int_sex_trestbps$`Resid. Dev`[2],
    int_sex_exang$`Resid. Dev`[2],
    int_sex_slope$`Resid. Dev`[2],
    int_sex_ca$`Resid. Dev`[2],
    int_cp_trestbps$`Resid. Dev`[2],
    int_cp_exang$`Resid. Dev`[2],
    int_cp_slope$`Resid. Dev`[2],
    int_cp_ca$`Resid. Dev`[2],
    int_trestbps_exang$`Resid. Dev`[2],
    int_trestbps_slope$`Resid. Dev`[2],
    int_trestbps_ca$`Resid. Dev`[2],
    int_exang_slope$`Resid. Dev`[2],
    int_exang_ca$`Resid. Dev`[2],
    int_slope_ca$`Resid. Dev`[2]
)

df4 = c(
    int_sex_cp$`Resid. Df`[2],
    int_sex_trestbps$`Resid. Df`[2],
    int_sex_exang$`Resid. Df`[2],
    int_sex_slope$`Resid. Df`[2],
    int_sex_ca$`Resid. Df`[2],
    int_cp_trestbps$`Resid. Df`[2],
    int_cp_exang$`Resid. Df`[2],
    int_cp_slope$`Resid. Df`[2],
    int_cp_ca$`Resid. Df`[2],
    int_trestbps_exang$`Resid. Df`[2],
    int_trestbps_slope$`Resid. Df`[2],
    int_trestbps_ca$`Resid. Df`[2],
    int_exang_slope$`Resid. Df`[2],
    int_exang_ca$`Resid. Df`[2],
    int_slope_ca$`Resid. Df`[2]
)

dfDifference4 = c(
    int_sex_cp$Df[2],
    int_sex_trestbps$Df[2],
    int_sex_exang$Df[2],
    int_sex_slope$Df[2],
    int_sex_ca$Df[2],
    int_cp_trestbps$Df[2],
    int_cp_exang$Df[2],
    int_cp_slope$Df[2],
    int_cp_ca$Df[2],
    int_trestbps_exang$Df[2],
    int_trestbps_slope$Df[2],
    int_trestbps_ca$Df[2],
    int_exang_slope$Df[2],
    int_exang_ca$Df[2],
    int_slope_ca$Df[2]    
)

AIC4 = c(
    model_se.cp.tr.ex.sl.ca_se.cp$aic,
    model_se.cp.tr.ex.sl.ca_se.tr$aic,
    model_se.cp.tr.ex.sl.ca_se.ex$aic,
    model_se.cp.tr.ex.sl.ca_se.sl$aic,
    model_se.cp.tr.ex.sl.ca_se.ca$aic,
    model_se.cp.tr.ex.sl.ca_cp.tr$aic,
    model_se.cp.tr.ex.sl.ca_cp.ex$aic,
    model_se.cp.tr.ex.sl.ca_cp.sl$aic,
    model_se.cp.tr.ex.sl.ca_cp.ca$aic,
    model_se.cp.tr.ex.sl.ca_tr.ex$aic,
    model_se.cp.tr.ex.sl.ca_tr.sl$aic,
    model_se.cp.tr.ex.sl.ca_tr.ca$aic,
    model_se.cp.tr.ex.sl.ca_ex.sl$aic,
    model_se.cp.tr.ex.sl.ca_ex.ca$aic,
    model_se.cp.tr.ex.sl.ca_sl.ca$aic
)

DevianceDifference4 = c(
    int_sex_cp$Deviance[2],
    int_sex_trestbps$Deviance[2],
    int_sex_exang$Deviance[2],
    int_sex_slope$Deviance[2],
    int_sex_ca$Deviance[2],
    int_cp_trestbps$Deviance[2],
    int_cp_exang$Deviance[2],
    int_cp_slope$Deviance[2],
    int_cp_ca$Deviance[2],
    int_trestbps_exang$Deviance[2],
    int_trestbps_slope$Deviance[2],
    int_trestbps_ca$Deviance[2],
    int_exang_slope$Deviance[2],
    int_exang_ca$Deviance[2],
    int_slope_ca$Deviance[2]    
)
Pvalue4 = c(
    int_sex_cp$`Pr(>Chi)`[2],
    int_sex_trestbps$`Pr(>Chi)`[2],
    int_sex_exang$`Pr(>Chi)`[2],
    int_sex_slope$`Pr(>Chi)`[2],
    int_sex_ca$`Pr(>Chi)`[2],
    int_cp_trestbps$`Pr(>Chi)`[2],
    int_cp_exang$`Pr(>Chi)`[2],
    int_cp_slope$`Pr(>Chi)`[2],
    int_cp_ca$`Pr(>Chi)`[2],
    int_trestbps_exang$`Pr(>Chi)`[2],
    int_trestbps_slope$`Pr(>Chi)`[2],
    int_trestbps_ca$`Pr(>Chi)`[2],
    int_exang_slope$`Pr(>Chi)`[2],
    int_exang_ca$`Pr(>Chi)`[2],
    int_slope_ca$`Pr(>Chi)`[2]
)

SIGNIFICANT4 = c(
    ifelse(int_sex_cp$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_sex_trestbps$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_sex_exang$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_sex_slope$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_sex_ca$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_cp_trestbps$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_cp_exang$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_cp_slope$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_cp_ca$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_trestbps_exang$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_trestbps_slope$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_trestbps_ca$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_exang_slope$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_exang_ca$`Pr(>Chi)`[2]< alpha4,'YES','NO'),
    ifelse(int_slope_ca$`Pr(>Chi)`[2]< alpha4,'YES','NO')
)

                  
ModelsCompared4 = c(
    '(27)-(25)',
    '(28)-(25)',
    '(29)-(25)',
    '(30)-(25)',
    '(31)-(25)',
    '(32)-(25)',
    '(33)-(25)',
    '(34)-(25)',
    '(35)-(25)',
    '(36)-(25)',
    '(37)-(25)',
    '(38)-(25)',
    '(39)-(25)',
    '(40)-(25)',
    '(41)-(25)'
)

step4 = data.frame(
    Model4,ExplanatoryVariables4,Deviance4,df4,AIC4,ModelsCompared4,DevianceDifference4,dfDifference4,Pvalue4,SIGNIFICANT4
)



```

```{r}
step4

```

## Purposeful Selection final model

```{r}
##useful model
modelPS = model_se.cp.tr.ex.sl.ca
summary(modelPS)

```

## AIC as categorical

```{r}
##Step wise: smaller AIC is better
completeModel <- glm(target ~ age+factor(sex)+factor(cp)+trestbps+chol+factor(fbs)
                                              +factor(restecg)+thalach+factor(exang)+oldpeak+factor(slope)+factor(ca), 
                                              family = binomial, data=heart)

# stepwise backward selection using AIC
library(MASS)
stepAIC(completeModel) 

```

### AIC model

```{r}
modelAIC <- glm(target ~ thalach+oldpeak+factor(exang)+trestbps+factor(slope)+factor(cp)+factor(sex)+factor(ca),
                family = binomial, data=heart)
summary(modelAIC)

```

## 3. BIC as numerical

```{r}
##bestglm with BIC
library(leaps)
library(bestglm)
bestglm(heart, family=binomial, IC="BIC")

```

### BIC model

```{r}
#modelBIC
modelBIC <- glm(target ~ sex+cp+thalach+exang+oldpeak+ca, family = binomial, data=heart)
summary(modelBIC)

```

## ROC curve

```{r}

library(pROC)
rocplot_1 <- roc(target ~ fitted(modelPS), data=heart)
rocplot_2 <- roc(target ~ fitted(modelAIC), data=heart)
rocplot_3 <- roc(target ~ fitted(modelBIC), data=heart)
plot.roc(rocplot_1, legacy.axes=TRUE, col ="green") 
plot.roc(rocplot_2, legacy.axes=TRUE, add = TRUE, col="purple") 
plot.roc(rocplot_3, legacy.axes=TRUE, add = TRUE, col="yellow") 
legend("bottomright",c('modelPS',
                   'modelAIC',
                   'modelBIC'), 
       lty=c(1, 1), col=c("green", "purple","yellow"), cex=0.7)

print('modelPS: ')
modelPS$formula
print('modelAIC: ')
modelAIC$formula
print('modelBIC: ')
modelBIC$formula

data.frame(modelPS = auc(rocplot_1),
           modelAIC = auc(rocplot_2),
           modelBIC = auc(rocplot_3))
```

## AIC values

```{r}
data.frame(modelSS = AIC(modelPS),
           modelAIC = AIC(modelAIC),
           modelBIC = AIC(modelBIC))

```

## Summarizing Predictive Power: Multiple Correlation (R)

```{r}

R1 = cor(target, fitted(modelPS))
R2 = cor(target, fitted(modelAIC))
R3 = cor(target, fitted(modelBIC))

data.frame(modelPS = R1,
           modelAIC = R2,
           modelBIC = R3)


```
